---
title: "Search Options"
output:
  html_document:
    includes: null
---

By default, searching does not use regular expressions, and is case-insensitive. You can turn on regular expressions by using the option `search = list(regex = TRUE)`, and perform case-sensitive searching by `list(search = caseInsensitive = FALSE)` (the two options are orthogonal).

```{r}
library(DT)
mtcars2 = mtcars[, c(1:5, 9)]
mtcars2$am = factor(mtcars$am, c(0, 1), c('automatic', 'manual'))
# search for Ma or Me
datatable(
  mtcars2, colnames = c('model' = 1),
  filter = list(position = 'top', clear = FALSE),
  options = list(
    search = list(regex = TRUE, caseInsensitive = FALSE, search = 'M[ae]'),
    pageLength = 5
  )
)
```

Please note the interpretation of regular expressions depends on whether you use client-side or server-side processing. For client-side processing, the search string is treated as JavaScript regular expressions. For server-side processing, it is up to the server-side script. In **DT**, the server-side processing function is implemented in R, so the search string is treated as R regular expressions.

In the above example, we specified an initial search string `M[ae]` (meaning `Ma` or `Me`) in the global search box. We can also specify initial search strings for individual columns via the `searchCols` option when column filters are enabled. The `searchCols` option is a list of the same length as the number of columns of the table. Each element in the list is either `NULL` or a list of the form `list(search = 'keyword')`. In the example below, we search for `er` in the first column, filter the second column using `[15, 25]`, and filter the last column by `"automatic"`.

```{r}
datatable(
  mtcars2, colnames = c('model' = 1),
  filter = list(position = 'top', clear = FALSE),
  options = list(pageLength = 5, searchCols = list(
    list(search = 'er'),
    list(search = '15 ... 25'),
    NULL, NULL, NULL, NULL,
    list(search = '["automatic"]')
  ))
)
```

