---
title: Helper Functions
---

**DT** has provided a few helper functions to make it convenient to tweak the table without writing a lot of HTML or JavaScript code.

---

# Format Columns

You can use the functions `format*()` to format the table columns as currency, percentages, or round numbers. Here are some examples:

```{r rand.seed=set.seed(123)}
m = cbind(matrix(rnorm(60, 1e5, 1e6), 20), runif(20), rnorm(20, 100))
m[, 1:3] = round(m[, 1:3])
m[, 4:5] = round(m[, 4:5], 7)
colnames(m) = head(LETTERS, ncol(m))
head(m)

# format the columns A and C as currency, and D as percentages
datatable(m) %>% formatCurrency(c('A', 'C')) %>% formatPercentage('D', 2)

# the first two columns are Euro currency, and round column E to 3 decimal places
datatable(m) %>% formatCurrency(1:2, '\U20AC') %>% formatRound('E', 3)
```

Under the hood, these formatting functions are just wrappers for the `rowCallback` option to generate the appropriate JavaScript code.

# Table Utilities

The functions `tableHeader()` and `tableFooter` are wrappers of `htmltools::tags` to generate a table header or footer. It may be convenient when you want to define a custom [table container](./).

```{r comment='', render=print}
tableHeader(iris)
tableHeader(colnames(iris))  # equivalent to tableHeader(iris)
tableFooter(iris)  # footer

library(htmltools)
tags$table(tableHeader(iris), tableFooter(iris))
```
