---
title: Helper Functions
---

**DT** has provided a few helper functions to make it convenient to tweak the table without writing a lot of HTML or JavaScript code.

---

# Format Columns

You can use the functions `format*()` to format the table columns as currency, percentages, or round numbers. Here are some examples:

```{r rand.seed=set.seed(123)}
m = cbind(matrix(rnorm(60, 1e5, 1e6), 20), runif(20), rnorm(20, 100))
m[, 1:3] = round(m[, 1:3])
m[, 4:5] = round(m[, 4:5], 7)
colnames(m) = head(LETTERS, ncol(m))
head(m)

# format the columns A and C as currency, and D as percentages
datatable(m) %>% formatCurrency(c('A', 'C')) %>% formatPercentage('D', 2)

# the first two columns are Euro currency, and round column E to 3 decimal places
datatable(m) %>% formatCurrency(1:2, '\U20AC') %>% formatRound('E', 3)
```

Under the hood, these formatting functions are just wrappers for the `rowCallback` option to generate the appropriate JavaScript code.

Similarly, there is a `formatDate()` function that can be used to format date/time columns. It has a `method` argument that takes values from a list of possible [conversion methods](http://mzl.la/1xGe99W): `r paste(sprintf('\x60%s\x60', DT:::DateMethods), collapse = ', ')`.

```{r collapse=TRUE}
x = strptime('2015-03-27 14:12:30', '%Y-%m-%d %H:%M:%S', tz = 'CDT')
(x = x + seq(1, 1e6, length.out = 10))
DT:::DateMethods
# use m = DT:::DateMethods below to see the output of all methods
m = c('toDateString', 'toLocaleDateString', 'toLocaleString', 'toUTCString')
d = as.data.frame(setNames(lapply(m, function(.) x), m))
str(d)
datatable(d, options = list(pageLength = 5, dom = 'tip')) %>%
  formatDate(1:ncol(d), m)
```

Please note that these conversion methods may be dependent on your system locale and web browser. The same method may lead to different output in different web browsers and system languages.

# Table Utilities

The functions `tableHeader()` and `tableFooter` are wrappers of `htmltools::tags` to generate a table header or footer. It may be convenient when you want to define a custom [table container](./).

```{r comment='', render=print}
tableHeader(iris)
tableHeader(colnames(iris))  # equivalent to tableHeader(iris)
tableFooter(iris)  # footer

library(htmltools)
tags$table(tableHeader(iris), tableFooter(iris))
```
