---
title: DataTables Plug-ins
---

---

**DataTables** has provided a large number of [plug-ins](http://datatables.net/plug-ins/), and we have integrated a subset of them in **DT**. You may enable plug-ins via the `plugins` argument of `datatable()`, e.g. `plugins = c('natural')`. If you find any plug-ins that you are interested in but are not listed here, you may file a feature request using [Github issues](https://github.com/rstudio/DT/issues). Currently these plug-ins have been integrated into **DT**:

```{r, results='asis', echo=FALSE}
plugins_txt = sort(names(DT:::available_plugins()))
plugins = strsplit(plugins_txt, strsplit, split = '/', fixed = TRUE)
last_group = ''
for (plugin in plugins) {
  group = plugin[1]
  value = plugin[2]
  if (group != last_group) {
    cat('- ', group, '\n')
    last_group = group
  }
  cat('    - ', value, '\n')
}
```

Please note the plug-ins for filtering and sorting may not work for server-side processing, e.g. when you render tables in Shiny.

# Examples

Here is an example of the plug-in `natural.js` for natural sorting. Compare the last column of the two tables after sorting:

```{r}
library(DT)
d = cbind(
  head(iris),
  Mixed = c(1, 2, 3, '10', '1a', '1b')
)
datatable(d, options = list(dom = 't'))
datatable(d, plugins = 'natural', options = list(
  dom = 't',
  columnDefs = list(list(type = 'natural', targets = 6))
))
```

