# A Minimal Example

Here is a minimal example of (client-side) DataTables in Shiny using the convenience functions `dataTableOutput()` and `renderDataTable()` in **DT** (instead of **shiny**):

```{r eval=FALSE}
library(shiny)
shinyApp(
  ui = fluidPage(DT::dataTableOutput('tbl')),
  server = function(input, output) {
    output$tbl = DT::renderDataTable({
      DT::datatable(iris)
    })
  }
)
```

If you have used the `renderDataTable()` function in **shiny** before, please note that the function with the same name in **DT** should take a widget as the input instead of a data object, i.e. you should `renderDataTable(datatable(iris))` instead of `renderDataTable(iris)` in **DT**.

# Interaction with Shiny

There are some information exposed to Shiny from the table widget as you interact with the table in Shiny.

## Row Selection

One way to enable row selection in the table is to add a column of checkboxes to the table. You can use the function `appendCheckboxes()` to do this. Remember, you need to tell `datatable()` not to escape this column, otherwise you will see raw HTML code like `<input type="checkbox" ... />` in the table.

```{r}
iris2 = head(iris, 20)
datatable(
  appendCheckboxes(iris2), escape = -7,
  options = list(pageLength = 5, dom = 'tip')
)
# just use checkboxes as the "row names"
datatable(
  iris2, rownames = checkboxRows(iris2), escape = -1,
  options = list(pageLength = 5, dom = 'tip')
)
```

The examples above only show you how a table looks like when you add a column of checkboxes. See [this example](https://yihui.shinyapps.io/DT-checkbox/) for how Shiny react to changes from the checkboxes.
